Hypertension and microbiome

### 2018-12-31 rrbiome.Rmd
Group comparison

```{r data}
pseq <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
```

@@

@@

```{r filtering}
dset <- meta(pseq) %>%
    tibble::rownames_to_column(var = "rowname") %>%
    dplyr::select(rowname, DNA_OK, MEN, BL_AGE, SYSTM, DIASM, BP_TREAT, BMI, CURR_SMOKE, DIAB, BL_USE_RX_C09,
        Q57X, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
    dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
                  HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
                  SEX = factor(ifelse(MEN == "Female", 1, 0)),
                  MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
    dplyr::select(-MEN) %>%
    stats::na.omit() %>%
    tibble::remove_rownames() %>%
    tibble::column_to_rownames(var = "rowname")

print.dset(dset)
```

Continuous variables

```{r distribution plots, fig.width=10, fig.height=4}
ggplot((dset %>%
        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP) %>%
        gather(Covariate, Value)), aes_string(x = "Value")) +
    facet_wrap(~Covariate, ncol=4, scales = "free") +
    geom_density(aes(y=..scaled..)) + 
    ggplot2::theme_minimal() 
```

Categorical variables

```{r histogramsy, fig.width=10, fig.height=4}
ggplot((dset %>%
        select(-BL_AGE, -BMI, -SYSTM, -DIASM, -PULSEPRESSURE, -MAP) %>%
        gather(Covariate, Value)), aes_string(x = "Value")) +
    facet_wrap(~Covariate, ncol=8, scales = "free") +
    geom_bar() + 
    ggplot2::theme_minimal() 
```

```{r Updating meta data}
pseq_filtered <- pseq
sample_data(pseq_filtered) <- sample_data(dset)
```

>>>>

filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>%
        dplyr::mutate(PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | BP_TREAT == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
 #       dplyr::filter(DNA_OK == 1) %>%
        stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")
```

@@

```{r data}
pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
```

@@

```{r filtering}
dset.genus <- filter.phenotype.data(pseq.genus)
dset.species <- filter.phenotype.data(pseq.species)
```

Phenotypical data between genus and species all.equal `r
isTRUE(all.equal(dset.genus, dset.species))`. Dimension for data
frames are `r dim(dset.genus)` and `r dim(dset.species)`.

Updating phyloseq objects

```{r updating phyloseq objects}
pseq.genus <- sample_data(dset.genus)
pseq.species <- sample_data(dset.species)
```

# How large subgroups are?

Number of participants in each combination of categorical variables

```{r numbers by group}
dset.species %>%
    group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION, Q57X,
             BL_USE_RX_C09, BL_USE_RX_C03, BL_USE_RX_C07, BL_USE_RX_C08) %>%
    select(-SYSTM, -DIASM, -BL_AGE, -BMI, -PULSEPRESSURE, -MAP) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable            
```

Number of particiapants in each combination of subset of categorical variables

```{r numbers by group reduced}
dset.species %>% group_by(SEX, BP_TREAT, CURR_SMOKE,  HYPERTENSION) %>%
    select(SEX, BP_TREAT, CURR_SMOKE, HYPERTENSION) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    kable
```
